---
title: "estimating group means with bootstrapped"
author: "Timo van Ommeren"
date: "2026-03-11"
output: html_document
---

```{r}
library(here)
here::i_am("Analysis/analysis/main_analysis.Rmd")

source(here("Analysis/R/01_load_data.R"))
source(here("Analysis/R/02_prepare_data.R"))
source(here("Analysis/R/03_descriptives.R"))
# source(here("R/04_models.R"))
# source(here("R/05_plots.R"))

```

# Loading and transforming data

```{r load and transform data}

# import simulation results and metadata (specify paths relative to project root)
raw <- load_simulation_data(data = "simulation_results/run_01/all_simulation_results.csv", 
                             metadata = "Analysis/percentage_relevant.csv")

# (1) transform simulation data to wide
# (2) add the number of records, percentage of relevant records and dataset topic(s) from metadata to simulation data (these will serve as group-level predictors in multilevel model)
# (4) rescale n_abstracts (divide by 100) and length_abstracts (divide by 1000)
# (5) truncate topic names to first topic only
processed <- prepare_data(raw$simulation, raw$meta)



processed$simulation <- processed$simulation %>%
  filter(run <= 270)

processed$simulation <- processed$simulation %>%
  filter(dataset != "Moran_2021")

# two-part coding of llm-specific predictors (implicit dummy variable creation through zero imputation, see reproduction_Dziak_Henry.Rmd for verification of method)
processed$simulation <- two_part_coding(processed$simulation)

# assign to variables for easier access
simulation_long <- raw$simulation
simulation      <- processed$simulation
metadata        <- processed$meta

```

# Preparation for analysis

```{r set colours of plots}

cud_cols <- c(
  llm               = "#009E73",  # bluish green
  criteria          = "#E69F00",  # orange
  random            = "#0072B2",  # blue
  no_initialisation = "#D55E00"   # vermillion
)


vir_cols <- c(
    llm=                "#482475",
    criteria=           "#2d708e",
    random=             "#2ab07f",
    no_initialisation=  "#bddf26"
)

# vir_cols <- setNames(
#   viridisLite::viridis(4, option = "viridis", begin = 0.10, end = 0.90),
#   c("llm", "criteria", "random", "no_initialisation")
# )

#define colours so that llm is green, criteria is yellow, no initialisation is red and random is blue

cols <- c(
  "llm" = "green", 
  "criteria" = "yellow", 
  "random" = "blue", 
  "no_initialisation" = "red"
)

```


## Descriptives

### 

```{r descriptive table}
descriptive_tables(simulation)

#export descriptive statistics table to latex format text file
writeLines(capture.output(descriptive_tables(simulation)), here::here("Report/results/descriptives_table.tex"))
```


### Descriptive plots

```{r barchart}
descriptive_barchart(simulation, metadata, papers_found, y_axis = 100, colours = vir_cols)
```

Conclusion from plot: dataset should be used as control variable.

```{r}
descriptive_histogram(simulation, colours = vir_cols)
```
Conclusion from plot: if starting performance is modelled as a binomial variance, we may expect overdispersion (i.e., a larger variance than would be expected from the model-derived variance, given the mean-variance relationship of the binomial). Histograms per dataset show that even when controlling for dataset this is the case. note for example the spike at zero.

```{r homogenity test}

check_homogeneity(simulation, colours = vir_cols)

#return the variance of papers found by condition
simulation %>% group_by(condition) %>%
  summarise(
    variance = var(papers_found),
    mean = mean(papers_found)
  )

```


```{r}
simulation <- simulation %>% mutate(condition = relevel(factor(condition), ref = "llm"))
```


## Model

```{r}
OLS_model <- lm(papers_found ~ condition - 1 + dataset + n_trials, 
                data = simulation)

summary(OLS_model)
```


