---
title: "main_analysis"
author: "Timo van Ommeren"
date: "2026-01-22"
output: html_document
---

```{r}
library(here)
here::i_am("Analysis/analysis/main_analysis.Rmd")

source(here("Analysis/R/01_load_data.R"))
source(here("Analysis/R/02_prepare_data.R"))
source(here("Analysis/R/03_descriptives.R"))
# source(here("R/04_models.R"))
# source(here("R/05_plots.R"))

```

## Loading and transforming data

```{r load and transform data}

# import simulation results and metadata (specify paths relative to project root)
raw <- load_simulation_data(data = "simulation_results/correct_trials/all_simulation_results.csv", 
                             metadata = "Analysis/percentage_relevant.csv")

# (1) transform simulation data to wide
# (2) add the number of records, percentage of relevant records and dataset topic(s) from metadata to simulation data (these will serve as group-level predictors in multilevel model)
# (4) rescale n_abstracts (divide by 100) and length_abstracts (divide by 1000)
# (5) truncate topic names to first topic only
processed <- prepare_data(raw$simulation, raw$meta)



processed$simulation <- processed$simulation %>%
  filter(run <= 270)

processed$simulation <- processed$simulation %>%
  filter(dataset != "Moran_2021")

# two-part coding of llm-specific predictors (implicit dummy variable creation through zero imputation, see reproduction_Dziak_Henry.Rmd for verification of method)
processed$simulation <- two_part_coding(processed$simulation)

# assign to variables for easier access
simulation_long <- raw$simulation
simulation      <- processed$simulation
metadata        <- processed$meta

```


## Descriptives

```{r}
descriptive_barchart(simulation, metadata, papers_found, y_axis = 100)
```


```{r}
ggplot(simulation, aes(x = papers_found / n_trials, fill = condition)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(
    title = "Distribution of number of relevant records found by condition",
    x = "Relevant records found (X) divided by number of trials (n <=100)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 20, 10, 10)
  )
```

```{r}
# 1) Build a list of plots (one per dataset)
dataset_names <- unique(simulation$dataset)

plots <- lapply(dataset_names, function(dataset_name) {
  ggplot(simulation %>% filter(dataset == dataset_name),
         aes(x = papers_found / n_trials, fill = condition)) +
    geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
    labs(
      title = paste("N relevant records found by condition -", dataset_name),
      x = "Relevant records found (X) divided by number of trials (n <=100)",
      y = "Count"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background  = element_rect(fill = "white", color = NA),
      legend.background = element_rect(fill = "white", color = NA),
      legend.key = element_rect(fill = "white", color = NA),
      plot.margin = margin(10, 20, 10, 10)
    )
})

# 2) Write to a multi-page PDF with 4 plots per page
pdf("..\\results\\papers_found_histograms.pdf", width = 11, height = 8.5, onefile = TRUE)

for (i in seq(1, length(plots), by = 4)) {
  page_plots <- plots[i:min(i + 3, length(plots))]
  print(patchwork::wrap_plots(page_plots, ncol = 2, nrow = 2))
}

dev.off()

```

```{r}
descriptive_histogram(simulation, papers_found / n_trials, split_by_condition = TRUE, bins = 100, save_path = here::here("Report/results/papers_found_histogram.png"), width = 10, height = 6, dpi = 300)
```

- NOTE TO ADD HISTOGRAM SPLIT BY DATASET TO PLOT ABOVE

- UNIQUE TABLE

## Data Analysis

First and foremost, we will set the experimental LLM condition as a reference level for all analyses. 

```{r}
simulation <- simulation %>% mutate(condition = relevel(factor(condition), ref = "llm"))
```


The analysis starts with a simple OLS analysis of the difference in number of papers found between conditions. For all subsequent models the marginal mean of papers found while controlling for the number of trials per condition will be estimated. As such, the primary OLS analysis can be seen as baseline to compare with the more complex multilevel models that follow.

```{r}

ols_model <- lm(papers_found / n_trials ~ condition, data = simulation)
summary(ols_model)


emmeans::emmeans(ols_model, ~ condition, type = "response")
```
```{r}
binomial_model <- glm(cbind(papers_found, n_trials - papers_found) ~ condition, data = simulation, family = binomial)

summary(binomial_model)


emmeans::emmeans(binomial_model, ~ condition, type = "response")
```
The estimated marginal means of the binomial model is near identical to the OLS model, indicating that the OLS model is sufficient for estimating the effect of condition on papers found.

```{r}
betabin_model <- glmmTMB::glmmTMB(
  cbind(papers_found, n_trials - papers_found) ~ condition + (1|dataset),
  family = glmmTMB::betabinomial(link = "logit"),
  dispformula = ~ 1,
  data = simulation
)

summary(betabin_model)

emmeans::emmeans(betabin_model, ~ condition, type = "response")
emmeans::emmeans(betabin_model, ~ condition, type=  "response", mode="latent")
```

